---
title: "Chad Huntebrinker's Homework 4"
author: "Chad Huntebrinker"
date: "2024-09-24"
output:
  word_document: default
  html_document: default
---

Question 2.4b and c: With a 99% confidence interval for B1, test using the test statistic t* to find a linear association.  Find the P-value.
```{r}
#Chad Huntebrinker
#Problem 2.4b
#Load the library and data in to R
library(readxl)

excel_data <- read_excel("Grade_Point_Average_Data.xlsx")

# GPA ~ ACT
# Fit the model and get the summary of the model
model_1 <- lm(GPA~ACT_Score,data=excel_data)
sum_of_model_1 <- summary(model_1)
sum_of_model_1
sum_of_model_1$coefficients[2,4]

# Plot the graph just to see
plot(GPA~ACT_Score,data=excel_data)
abline(model_1,col="red")

#Calculate the t score
b1 <- sum_of_model_1$coefficients[2, 1]
se_b1 <- sum_of_model_1$coefficients[2, 2]

#a = 0.01
t_1 <- b1 / se_b1

#If the t_1 > than the t score, than we conclude that H1 is true (which is B1 != 0)
t_1 > qt(0.995, 118)

#Problem 2.4c
#Calculate the p_value
p_1 <- 2 * (1 - pt(t_1, 118))

#If the p score is less than alpha, than H1 is true (which is B1 != 0)
p_1 < 0.01
```

Problem 2.6b: With a level of significance of .05, conduct a t test and find the P-value to see if there is a linear association
```{r}
#Chad Huntebrinker
#Problem 2.6b
#Load the library and data in to R
library(readxl)

excel_data <- read_excel("Airfreight_Breakage_Data.xlsx")

#Number_Broker ~ Number_Transferred
#Fit the model
model_2 <- lm(Number_Broker~Number_Transferred,data=excel_data)
sum_of_model_2 <- summary(model_2)
sum_of_model_2
sum_of_model_2$coefficients[2,4]

#Plot to see the graph
plot(Number_Broker~Number_Transferred, data = excel_data)
abline(model_2,col="blue")


#Calculate the t score
b1 <- sum_of_model_2$coefficients[2, 1]
se_b1 <- sum_of_model_2$coefficients[2, 2]

#a = 0.05
t_1 <- b1 / se_b1

#If the t_1 > than the t score, than we conclude that H1 is true (which is B1 != 0)
t_1 > qt(0.975, 8)

#Calculate the p_value
p_1 <- 2 * (1 - pt(t_1, 8))

#If the p score is less than alpha, than H1 is true (which is B1 != 0)
p_1 < 0.01

```

Problem 2.14a: Find the confidence interval with a 90% confidence interval for the mean service time on calls in which six copier are serviced
```{r}
#Chad Huntebrinker
#Problem 2.14a
#Load the library and data in to R
library(readxl)

excel_data <- read_excel("Copier_Maintenance_Data.xlsx")

#Total_Minutes ~ Number_Of_Copiers
#Fit the model
model_3 <- lm(Total_Minutes~Number_Of_Copiers,data=excel_data)
sum_of_model_3 <- summary(model_3)

#Plot to see the graph
plot(Total_Minutes~Number_Of_Copiers, data = excel_data)
abline(model_3,col="blue")


#Get b1 and se_b1
b1 <- sum_of_model_3$coefficients[2, 1]
se_b1 <- sum_of_model_3$coefficients[2, 2]

#Get the mean service time for 6 copiers serviced
Y_hat <- predict(model_3, data.frame(Number_Of_Copiers=6))

#Now get the confidence interval
predict(model_3, data.frame(Number_Of_Copiers=6), interval = "confidence", level = 0.90)



```

Problem 2.27a and c: With a control risk for type 1 error of .05, find if there is a negative linear association.  Estimate with a 95% confidence interval the difference in expected mm for women whose ages differ by one year.
```{r}
#Chad Huntebrinker
#Problem 2.27a
#Load the library and data in to R
library(readxl)

excel_data <- read_excel("Muscle_Mass_Data.xlsx")

#Muscle_Mass ~ Age
#Fit the model
model_4 <- lm(Muscle_Mass~Age,data=excel_data)
sum_of_model_4 <- summary(model_4)

#Plot to see the graph
plot(excel_data$Muscle_Mass~excel_data$Age)
abline(model_4,col="green")


#Get b1 and se_b1
b1 <- summary(model_4)$coefficients[2, 1]
se_b1 <- summary(model_4)$coefficients[2, 2]

#Find t score
t_1 <- b1 / se_b1

#If the t_1 > than the t score, than we conclude that H1 is true (which is B1 != 0)
t_1 < qt(0.05, 58)

#Find p score
p_1 <- pt(t_1, 58)

#Problem 2.27c
qt(0.975, 58)
b1
se_b1

#Lower confidence
b1 - (qt(0.975, 58) * se_b1)

#Higher confidence
b1 + (qt(0.975, 58) * se_b1)

```


Problem 2.65:Obtain a separate interval estimate of B1 for each egion.  Use a 95% confidence coefficient in each case.
```{r}
#Chad Huntebrinker
#Problem 2.65
#Load the library and data in to R
library(readxl)

excel_data <- read_excel("SENIC_Data.xlsx")

#First
#For Infection_Risk, plot the Data
plot(Length_of_Stay ~ Infection_Risk, data = excel_data, pch = 16, xlab = "Infection Risk", ylab = "Length of Stay")

#Fit a Linear Regression model
model_7 <- lm(Length_of_Stay ~ Infection_Risk, data = excel_data)

#Find b coefficients
b7_coefficients <- model_7$coefficients

#Add the fitted line
abline(b7_coefficients, lwd =2, lty =2, col="red")

#Second
#For Available_facilities_and_services, plot the Data
plot(Length_of_Stay ~ Available_facilities_and_services, data = excel_data, pch = 16, xlab = "Available facilities and services", ylab = "Length of Stay")

#Fit a Linear Regression model
model_8 <- lm(Length_of_Stay ~ Available_facilities_and_services, data = excel_data)

#Find b coefficients
b8_coefficients <- model_8$coefficients

#Add the fitted line
abline(b8_coefficients, lwd =2, lty =2, col="green")

#Third
#For Routine_Chest_X-ray, plot the Data
plot(Length_of_Stay ~ `Routine_Chest_X-ray`, data = excel_data, pch = 16, xlab = "Routine Chest X-ray", ylab = "Length of Stay")

#Fit a Linear Regression model
model_9 <- lm(Length_of_Stay ~ `Routine_Chest_X-ray`, data = excel_data)

#Find b coefficients
b9_coefficients <- model_9$coefficients

#Add the fitted line
abline(b9_coefficients, lwd =2, lty =2, col="blue")

#Get the summary of each model
sum7 <- summary(model_7)
sum8 <- summary(model_8)
sum9 <- summary(model_9)
sum7$coefficients[2,4]
sum8$coefficients[2,4]
sum9$coefficients[2,4]

#Calculate the t score
b7 <- summary(model_7)$coefficients[2, 1]
se_b7 <- summary(model_7)$coefficients[2, 2]

b8 <- summary(model_8)$coefficients[2, 1]
se_b8 <- summary(model_8)$coefficients[2, 2]

b9 <- summary(model_9)$coefficients[2, 1]
se_b9 <- summary(model_9)$coefficients[2, 2]

#a = 0.05
t_7 <- b7 / se_b7
t_8 <- b8 / se_b8
t_9 <- b9 / se_b9

#Calculate the lower and upper confidence bounds for each variable
b7 - (qt(0.975, 112) * se_b7)
b7 + (qt(0.975, 112) * se_b7)

b8 - (qt(0.975, 112) * se_b8)
b8 + (qt(0.975, 112) * se_b8)

b9 - (qt(0.975, 112) * se_b9)
b9 + (qt(0.975, 112) * se_b9)

```

